---
title: "ESM 206 Lab 5"
author: "Allison Hacker"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


### Lab 5 objectives
- Learn to parse dates with "lubridate"
- Get counts of observations with count()
- Uncount() rows
- One and two sample t-test
- Create a heatmap with geom_tile()

```{r, include=FALSE}

#read in packages and attach data
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"), na = "-99999") %>% 
  clean_names() 
  
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```

### Exploratory data visualization

We're only going to consider site as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm))+
  geom_jitter(aes(color = site), alpha = 0.5, width = 0.2)

ggplot(lobster_tidy, aes(x = size_mm))+
  geom_histogram(aes(fill = site))+
  facet_wrap(~site) # can also use scales = "free" as an argument to facet wrap which will make the scales different between graphs

ggplot(lobster_tidy, aes(sample = size_mm))+
  geom_qq()+
  facet_wrap(~site)

```

### Convert the date column to class "Date"

We'll use lubridate package to convert to Date format, and then to help us easily parse month and year

```{r}
lobster_date <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))
```

Now let's parse year and month using lubridate::month() and lubridate::year()

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Now let's find counts of observed lobsters based on different hierarchical groupings:

First, let's count lobsters by year and month.
```{r}

lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y<- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_site <- lobster_parse_date %>% 
  count(site, obs_year)

```

If we want to create a summary table that contains multiple statistics other than count by group, it's easier to use group_by() and n()

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>%
  summarize( 
            mean_size = mean(size_mm, na.rm = TRUE),
            sd_size = sd(size_mm, na.rm = TRUE),
            lobster_number = n()
            )

```

### Find confidence intervals

Use t.test() function to find confidence intervals for one sample and t-test to compare means of two samples

```{r}
ivee_lobster <- lobster_tidy %>%
  filter(site == "IVEE") %>% 
  pull(size_mm) #creates a vector from the dataframe

#Find confidence interval

t.test(ivee_lobster)

```

### Two-sample t-test to compare means

We are basically asking if two samples are drawn from populations with the same mean

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs.

* We have already done our necessary exploratory analyses to determine that a two-sample t-test for means comparison is appropriate.

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest

```
There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p <0.001 , alpha = 0.5). 

^ This should be the least interesting part of your data analysis.

An easier way to do this:
```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("MOHK", "NAPL"))

mn_ttest2 <- t.test(data = lobster_mn, size_mm ~ site)
mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic
```

We can use inline variable referencing in R markdown to call a variable in text. This is better for reproducibility because it will update if data is changed.

Here is the p-value: `r mn_ttest2$p.value`

Here is the t-statistic: `r mn_ttest2$statistic`

### Heatmap!

```{r}

ggplot(data = lobster_site, aes(x = obs_year, y = site))+
  geom_tile(aes(fill = n))
```

